Paper_Results

### 2025-12-04 Paper_Results.Rmd
revision r1 AMPPS

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
library(dplyr)
library(lme4)
library(caret)
library(pROC)
library(patchwork)
```
## General Information
This R Markdown file can be used to replicate our results that are presented in the paper. Please note that it does not contain the shiny app.

@@

set.seed(12361488)
features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,0,0)
features_sample <- features_sample[[1]]

@@

table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
rownames(table) <- c("row-wise","subject-wise","rolling-window")
table <- round(table[,c(5,1,2,3,4)], 2)
table
```

@@

result <- read.csv("simulation_results_0.csv") # see Run_Simulation.R to investigate how results were created

p1 <- ggplot(result[result$A == 0,], aes(x=icc, y=auc, color = icc_pred)) + # color: icc_pred
  geom_point(alpha=0.4, size = 0.8) +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#1B4332", "#40916C", "#74C69D", "#B5E48C", "#FDE68A", "#FFB5A7")) +
  ylab("Prediction Performance (Overall AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome \n(ICC Outcome)") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  ggtitle("Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5") +
  annotate("text", x = 0.07, y = 0.82, label =  "Greater bias when between person differences explain a \nlarger % of total variation in both outcome and predictors.",  hjust = 0) +
  geom_curve(
    aes(x = 0.2, y = 0.86, xend = 0.35, yend = 0.93), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.2
  ) + 
  annotate("text", x = 0.38, y = 0.58, label = "Low bias if between person differences explain a small \n% of total variation in outcome, predictors, or both.",  hjust = 0) +
    geom_curve(
    aes(x = 0.37, y = 0.55, xend = 0.33, yend = 0.51), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.1
  ) 
  # Adding a diagonal line to illustrate the linear relationship
ggsave("Figures/Figure4_Example2.jpg",
  plot = p1,
  width = 10.2,
  height = 4.5)

@@

@@

  p = ggplot(df_long, aes(x = icc_data, y = AUC, color = AUC_type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess", span = spanpar) +
    theme_minimal() +
    scale_color_manual(
      values = c(
        "shuffle_AUC." = "#40916C",
        "movingwindowshuffle_AUC." = "#4D8FAC"
      ), 
      labels = c(
        "shuffle_AUC." = "No true Relationship (Record-Wise)",
        "movingwindowshuffle_AUC." = "No true Relationship (Moving-Window)"
      )
    ) +
    geom_hline(yintercept = 0.5, color = "grey", size = 0.8) +
    labs(color = "") +
    ylim(0.4, 0.95) +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),
                   ", Timepoints: ", max(data_viz$num_samples),
                   ",\nICC Predictors (mean): ", round(data_viz$icc_pred, 2))) +
    ylab("Overall AUC") +
    xlab("ICC Outcome")

@@

  p = ggplot(df_long, aes(x = icc_data, y = Correlation, color = type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess", span = spanpar) +
    theme_minimal() +
    scale_color_manual(values = c(
      "shuffle_cor" = "#40916C",
      "movingwindowshuffle_cor" = "#4D8FAC"
    ), labels = c("shuffle_cor" = "No true Relationship (Record-Wise)", 
                  "movingwindowshuffle_cor" ="No true Relationship (Moving-Window)"
    )) +
    labs(color = "") +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),
                   ", Timepoints: ", max(data_viz$num_samples),
                   ",\nICC Predictors (mean): ", round(data_viz$icc_pred, 2))) +
    ylab("Overall Correlation") +
    xlab("ICC Outcome") +
    ylim(0, 0.7)

@@

# Create the sorted plot with conditional background color
p2 <- ggplot(data, aes(x = time, y = y)) +
  # Add background shading for subjects with y = 1
  geom_rect(
    data = data %>% filter(!has_y1),
    aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
    fill = "#E5E4E2", alpha = 0.05, inherit.aes = T
  ) +
  geom_point(color = "#FC9D9A", size = 0.5) +                       
  facet_wrap(~subject) +               
  labs(
    x = "Time", 
    y = "Outcome", 
    title = paste("Within-person variability over time: \n78 participants (52%) do not experience within-person variability. ")
  ) +
  scale_y_continuous(breaks = c(0, 1)) +  # Ensure only 0 and 1 appear on the y-axis
  theme_minimal() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  guides(fill = "none")

@@

```{r }
############ Table 7 #############
##################################
sim_1$auc_value_base # row-wise
sim_4$auc_value_base # rolling window

@@

@@

plot_data = function(data_viz,spanpar){
  # Reshape to long format
  df_long <- data_viz %>%
    pivot_longer(
      cols = c(`real_AUC.random.intercept.only.`, `real_AUC.`,`movingwindow_AUC.`,`subjectwise_AUC.`,`movingwindow_AUC.random.intercept.only.`),
      names_to = "AUC_type",
      values_to = "AUC"
    )
  p = ggplot(df_long, aes(x = icc_data, y = AUC, color = AUC_type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess",span = spanpar) +
    theme_minimal() +
    scale_color_manual(values = c(
      "real_AUC.random.intercept.only." ="#B5E48C" ,
      "real_AUC." = "#74C69D",
      "movingwindow_AUC." ="#9EC5E7",
      "movingwindow_AUC.random.intercept.only." = "#A2D2FF",
      "subjectwise_AUC." ="#FFB5A7"
    ), 
    labels = c(
               "real_AUC.random.intercept.only."="Random Intercept only (Record-Wise)",
               "real_AUC." ="Random Forest (Record-Wise) ",
               "movingwindow_AUC." ="Random Forest (Moving-Window)",
               "subjectwise_AUC." ="Random Forest (Subject-Wise)",
               "movingwindow_AUC.random.intercept.only." = "Random Intercept only (Moving-Window)")) +
    theme(legend.position = "none") +
    geom_hline(yintercept = 0.5, color = "grey", size = 0.8) +
    labs(color = "Model") +
    ylim(0.4,0.95) +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),", ", "Timepoints: ", max(data_viz$num_samples),", ","\n",  "ICC Predictors (mean): ", round(data_viz$icc_pred,2))) +
    ylab("Overall AUC") +
    xlab("ICC Outcome")
  return(p)
}

@@

plot_data = function(data_viz,spanpar){
  # Reshape to long format
  df_long <- data_viz %>%
    pivot_longer(
    #  cols = c(`shuffle_mae`, `real_mae_base`, `real_mae`,`centered_mae`,`subjectwise_mae`),
      cols = c(`real_cor_base`, `real_cor`,`movingwindow_cor`,`subjectwise_cor`,movingwindow_cor_base),
      names_to = "type",
      values_to = "Correlation"
    )
  p = ggplot(df_long, aes(x = icc_data, y = Correlation, color = type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess",span = spanpar) +
    theme_minimal() +
    scale_color_manual(values = c(
      "real_cor_base" ="#B5E48C" ,
      "real_cor" = "#74C69D",
      "movingwindow_cor" = "#9EC5E7",
      "movingwindow_cor_base" = "#A2D2FF",
      "subjectwise_cor" ="#FFB5A7"
    ), labels = c("real_cor_base"="Random Intercept only (Record-Wise)",
                  "real_cor" ="Random Forest (Record-Wise)",
                  "movingwindow_cor" ="Random Forest (Moving-Window) ",
                  "subjectwise_cor" = "Random Forest (Subject-Wise)",
                  "movingwindow_cor_base" = "Random Intercept only (Moving-Window)"
    )) +
    theme(legend.position = "none") +
    labs(color = "Model") +
   # xlim(0.1,1) +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),", ", "Timepoints: ", max(data_viz$num_samples),", ","\n",  "ICC Predictors (mean): ", round(data_viz$icc_pred,2))) +
    ylab("Overall Correlation") +
    xlab("ICC Outcome") +
    ylim(0, 1)
  return(p)
}

@@

result = read.csv("simulation_results_baseline.csv")
p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
  geom_point(alpha=0.5, size = 1.2, color = "#40916C") +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#40916C")) +
  ylab("Overall Performance (AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  guides(col = guide_colourbar()) +
  ggtitle(paste("Performance of a Random Intercept Only Model (No Predictors Included)")) +
  geom_hline(yintercept = 0.5) +
  ylim(range= c(0.5,1)) +
    annotate("text", x = 0.5, y = 0.82, label =  "Performance increases as between person differences explain a \nlarger % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.7, y = 0.85, xend = 0.5, yend = 0.91), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "#1B4332", 
    curvature = 0.2
  ) +
      annotate("text", x = 0.2, y = 0.62, label =  "Performance can be moderate even if between person differences explain only a relatively \nsmall % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.19, y = 0.62, xend = 0.07, yend = 0.61), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "#1B4332", 
    curvature = 0.2
  )

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
library(dplyr)
library(lme4)
library(caret)
library(pROC)
library(patchwork)
library(CorBin)
```
## General Information
This R Markdown file can be used to replicate our results that are presented in the paper. Please note that it does not contain the shiny app.

@@

# Generating Features
A <- 0.05#1 # Relationship between features and outcome
feature_std <- 0.1 # population level feature generating process
B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
C <- 0.45 # Within-subject variability (added within participant for each timepoint)

source("Simulation_Functions_notshiny.R")
time_type = "None" #DCP, 1-dependent
time_effect = 0

set.seed(12361488)
features_sample <-create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, time_effect, time_type)
#features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,0,0)
features_sample <- features_sample[[1]]

@@

table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
rownames(table) <- c("row-wise","subject-wise","rolling-window")
table <- round(table[,c(5,1,2,3,4)], 2)
table

sub_id <- unique(features_sample$subject)

auc_value_ind  <- numeric(length(sub_id))
auc_value_base <- numeric(length(sub_id))
sub_number     <- sub_id  # initialize

for (k in seq_along(sub_id)) {

  idx <- features_sample$subject == sub_id[k]
  y_k <- features_sample$y[idx]

  # Only proceed if conditions are met
  if (sum(y_k) > 0 && length(unique(y_k)) > 1) {
    message("Processing subject ", k, " (ID: ", sub_id[k], ")")

    tryCatch(
      {
        res <- run_simulation_slidingwindow(
          features_sample[idx, ],
          1,
          windowsize = 14
        )

        auc_value_ind[k]  <- res$auc_value
        auc_value_base[k] <- res$auc_value_base
      },
      error = function(e) {
        warning("Simulation failed for subject ", sub_id[k], ": ", e$message)
        auc_value_ind[k]  <- NA
        auc_value_base[k] <- NA
      }
    )
  } else {
    auc_value_ind[k]  <- NA
    auc_value_base[k] <- NA
  }
}

mean(auc_value_ind, na.rm = TRUE)

sum(!is.na(auc_value_ind))

sum(auc_value_ind[!is.na(auc_value_ind)] != 0)

# Table A.1.1

time_type = "1-dependent" 
time_effect = 0.25

set.seed(12361488)
rm(features_sample)
features_sample <-create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, time_effect, time_type)
features_sample <- features_sample[[1]]

table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
rownames(table) <- c("row-wise","subject-wise","rolling-window")
table <- round(table[,c(5,1,2,3,4)], 2)
table

#Table A.1.2
time_type = "DCP"
time_effect =  seq(0.9, 0.01, length.out = (n_samples-1)) #0.2

set.seed(12361488)
rm(features_sample)
features_sample <-create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, time_effect, time_type)
features_sample <- features_sample[[1]]

table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
rownames(table) <- c("row-wise","subject-wise","rolling-window")
table <- round(table[,c(5,1,2,3,4)], 2)
table

# Table A.1.3
set.seed(12361488)
features_sample <-create_data(n_features, n_samples, n_subjects, A, feature_std = 2, B, C, overall_prob_outcome, sd_outcome, 0, "None")
features_sample <- features_sample[[1]]

table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
rownames(table) <- c("row-wise","subject-wise","rolling-window")
table <- round(table[,c(5,1,2,3,4)], 2)
table

@@

result <- read.csv("simulation_results_0.csv") # see Run_Simulation.R to investigate how results were created
title = "Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5"

p1 <- ggplot(result, aes(x=icc, y=auc, color = icc_pred)) + # color: icc_pred
  geom_point(alpha=0.4, size = 0.8) +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#1B4332", "#40916C", "#74C69D", "#B5E48C", "#FDE68A", "#FFB5A7")) +
  ylab("Prediction Performance (Overall AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome \n(ICC Outcome)") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  ggtitle(title) +
  annotate("text", x = 0.07, y = 0.82, label =  "Greater bias when between person differences explain a \nlarger % of total variation in both outcome and predictors.",  hjust = 0) +
  geom_curve(
    aes(x = 0.2, y = 0.86, xend = 0.35, yend = 0.93), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.2
  ) + 
  annotate("text", x = 0.38, y = 0.58, label = "Low bias if between person differences explain a small \n% of total variation in outcome, predictors, or both.",  hjust = 0) +
    geom_curve(
    aes(x = 0.37, y = 0.55, xend = 0.33, yend = 0.51), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.1
  ) 
#   # Adding a diagonal line to illustrate the linear relationship
# ggsave("Figures/Figure4_Example2.jpg",
#   plot = p1,
#   width = 10.2,
#   height = 4.5)

@@

result <- read.csv("simulation_results_time_A11.csv") # autoregressive effect in predictors
title = "Robustness Check 1: Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5"

result <- read.csv("simulation_results_time_A12.csv") # autoregressive effect in predictors
title = "Robustness Check 2: Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5"

result <- read.csv("simulation_results_time_A13.csv") # autoregressive effect in predictors
result <- result[result$feature_std ==2,]

title = "Robustness Check 3: Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5"

p1 <- ggplot(result, aes(x=icc, y=auc, color = icc_pred)) + # color: icc_pred
  geom_point(alpha=0.4, size = 0.8) +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#1B4332", "#40916C", "#74C69D", "#B5E48C", "#FDE68A", "#FFB5A7")) +
  ylab("Prediction Performance (Overall AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome \n(ICC Outcome)") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  ggtitle(title)

  # Adding a diagonal line to illustrate the linear relationship
ggsave("Figures/Figure_A13.jpg",
  plot = p1,
  width = 10.2,
  height = 4.5)

@@

  p = ggplot(df_long, aes(x = icc_data, y = AUC, color = AUC_type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess", span = spanpar) +
    theme_minimal() +
    scale_color_manual(
      values = c(
        "shuffle_AUC." = "#c7835a",
        "movingwindowshuffle_AUC." = "#58538f"
      ), 
      labels = c(
        "shuffle_AUC." = "No true Relationship (Record-Wise)",
        "movingwindowshuffle_AUC." = "No true Relationship (Moving-Window)"
      )
    ) +
    geom_hline(yintercept = 0.5, color = "grey", size = 0.8) +
    labs(color = "") +
    ylim(0.4, 0.95) +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),
                   ", Timepoints: ", max(data_viz$num_samples),
                   ",\nICC Predictors (mean): ", round(data_viz$icc_pred, 2))) +
    ylab("Overall AUC") +
    xlab("ICC Outcome")

@@

  p = ggplot(df_long, aes(x = icc_data, y = Correlation, color = type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess", span = spanpar) +
    theme_minimal() +
    scale_color_manual(values = c(
      "shuffle_cor" = "#c7835a",
      "movingwindowshuffle_cor" = "#58538f"
    ), labels = c("shuffle_cor" = "No true Relationship (Record-Wise)", 
                  "movingwindowshuffle_cor" ="No true Relationship (Moving-Window)"
    )) +
    labs(color = "") +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),
                   ", Timepoints: ", max(data_viz$num_samples),
                   ",\nICC Predictors (mean): ", round(data_viz$icc_pred, 2))) +
    ylab("Overall Correlation") +
    xlab("ICC Outcome") +
    ylim(-0.1, 0.7)

@@

# Create the sorted plot with conditional background color
p2 <- ggplot(data, aes(x = time, y = y)) +
  # Add background shading for subjects with y = 1
  geom_rect(
    data = data %>% filter(!has_y1),
    aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
    fill = "#E5E4E2", alpha = 0.05, inherit.aes = T
  ) +
  geom_point(color = "#c7835a", size = 0.5) +                       
  facet_wrap(~subject) +               
  labs(
    x = "Time", 
    y = "Outcome", 
    title = paste("Within-person variability over time: \n78 participants (52%) do not experience within-person variability. ")
  ) +
  scale_y_continuous(breaks = c(0, 1)) +  # Ensure only 0 and 1 appear on the y-axis
  theme_minimal() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  guides(fill = "none")

@@

```{r }
############ Table 7 #############
##################################
round(sim_1$auc_value_base,2) # row-wise
round(sim_4$auc_value_base,2) # rolling window
sim_3$auc_value_base # subject-wise

@@

test <- sim_3$ind_base[[1]] %>%
  group_by(subject) %>%
  filter(length(unique(true)) > 1) %>%
  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
  summarise(
    auc_val = pROC::auc(pROC::roc(as.numeric(as.character(true)), as.numeric(as.character(pred)), quiet = TRUE))[1],
    .groups = "drop" # Ungroup after summarizing
  )

@@

plot_data = function(data_viz,spanpar){
  # Reshape to long format
  df_long <- data_viz %>%
    pivot_longer(
      cols = c(`real_AUC.random.intercept.only.`, `real_AUC.`,`movingwindow_AUC.`,`subjectwise_AUC.`,`movingwindow_AUC.random.intercept.only.`),
      names_to = "AUC_type",
      values_to = "AUC"
    )
  p = ggplot(df_long, aes(x = icc_data, y = AUC, color = AUC_type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess",span = spanpar) +
    theme_minimal() +
    scale_color_manual(values = c(
      "real_AUC.random.intercept.only." ="#c7835a" ,
      "real_AUC." = "#9c5435",
      "movingwindow_AUC." ="#877fbf",
      "movingwindow_AUC.random.intercept.only." = "#58538f",
      "subjectwise_AUC." ="#40916c"
    ), 
    labels = c(
               "real_AUC.random.intercept.only."="Random Intercept only (Record-Wise)",
               "real_AUC." ="Random Forest (Record-Wise) ",
               "movingwindow_AUC." ="Random Forest (Moving-Window)",
               "subjectwise_AUC." ="Random Forest (Subject-Wise)",
               "movingwindow_AUC.random.intercept.only." = "Random Intercept only (Moving-Window)")) +
    theme(legend.position = "none") +
    geom_hline(yintercept = 0.5, color = "grey", size = 0.8) +
    labs(color = "Model") +
    ylim(0.4,0.95) +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),", ", "Timepoints: ", max(data_viz$num_samples),", ","\n",  "ICC Predictors (mean): ", round(data_viz$icc_pred,2))) +
    ylab("Overall AUC") +
    xlab("ICC Outcome")
  return(p)
}

@@

plot_data = function(data_viz,spanpar){
  # Reshape to long format
  df_long <- data_viz %>%
    pivot_longer(
    #  cols = c(`shuffle_mae`, `real_mae_base`, `real_mae`,`centered_mae`,`subjectwise_mae`),
      cols = c(`real_cor_base`, `real_cor`,`movingwindow_cor`,`subjectwise_cor`,movingwindow_cor_base),
      names_to = "type",
      values_to = "Correlation"
    )
  p = ggplot(df_long, aes(x = icc_data, y = Correlation, color = type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess",span = spanpar) +
    theme_minimal() +
    scale_color_manual(values = c(
      "real_cor_base" ="#c7835a" ,
      "real_cor" = "#9c5435",
      "movingwindow_cor" = "#877fbf",
      "movingwindow_cor_base" = "#58538f",
      "subjectwise_cor" ="#40916c"
    ), labels = c("real_cor_base"="Random Intercept only (Record-Wise)",
                  "real_cor" ="Random Forest (Record-Wise)",
                  "movingwindow_cor" ="Random Forest (Moving-Window) ",
                  "subjectwise_cor" = "Random Forest (Subject-Wise)",
                  "movingwindow_cor_base" = "Random Intercept only (Moving-Window)"
    )) +
    theme(legend.position = "none") +
    labs(color = "Model") +
   # xlim(0.1,1) +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),", ", "Timepoints: ", max(data_viz$num_samples),", ","\n",  "ICC Predictors (mean): ", round(data_viz$icc_pred,2))) +
    ylab("Overall Correlation") +
    xlab("ICC Outcome") +
    ylim(0, 1)
  return(p)
}

@@

result = read.csv("simulation_results_baseline.csv")
title = paste("Performance of a Random Intercept Only Model (No Predictors Included)")

result = read.csv("simulation_results_baseline_A21.csv")
title = paste("Robustness Check 1: Performance of a Random Intercept Only Model (No Predictors Included)")

result = read.csv("simulation_results_baseline_A22.csv")
title = paste("Robustness Check 2: Performance of a Random Intercept Only Model (No Predictors Included)")

p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
  geom_point(alpha=0.5, size = 1.2, color = "#40916C") +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#40916C")) +
  ylab("Overall Performance (AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  guides(col = guide_colourbar()) +
  ggtitle(title) +
  geom_hline(yintercept = 0.5) +
  ylim(range= c(0.5,1)) +
    annotate("text", x = 0.5, y = 0.82, label =  "Performance increases as between person differences explain a \nlarger % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.7, y = 0.85, xend = 0.5, yend = 0.91), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "#1B4332", 
    curvature = 0.2
  ) +
      annotate("text", x = 0.2, y = 0.62, label =  "Performance can be moderate even if between person differences explain only a relatively \nsmall % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.19, y = 0.62, xend = 0.07, yend = 0.61), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "#1B4332", 
    curvature = 0.2
  )

@@

ggsave(  "Figures/FigureA22.jpg",
  plot = p7,
  width = 10,
  height = 5)

@@

```{r }
# --- PARAMETERS ---
a_values <- c(0.00, 0.05)        # outcome effect
b_values <- c(0.1, 0.5, 0.9)     # cross-subject variability
c_values <- c(0.05, 0.25, 0.45)  # within-subject variability
d <- 0.1                         # noise

# --- COMPUTE SNR ---
results <- expand.grid(a = a_values, b = b_values, c = c_values)

results$var_signal <- results$a^2
results$var_noise  <- results$b^2 + results$c^2 + d^2
results$SNR_linear <- results$var_signal / results$var_noise
results$SNR_dB     <- ifelse(results$SNR_linear > 0,
                             10 * log10(results$SNR_linear),
                             -Inf)

# --- SHOW RESULTS ---
print(round(results, 6))

b_grid <- seq(0.1, 0.9, by = 0.05)
c_grid <- seq(0.05, 0.45, by = 0.05)
a <- 0.05

grid <- expand.grid(b = b_grid, c = c_grid)
grid$SNR_linear <- a^2 / (grid$b^2 + grid$c^2 + d^2)
grid$SNR_dB <- 10 * log10(grid$SNR_linear)

summary(grid$SNR_linear)
summary(grid$SNR_dB)

# Optional: visualize as heatmap
library(ggplot2)
ggplot(grid, aes(x = b, y = c, fill = SNR_linear)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "plasma") +
  labs(title = "SNR (dB) across b and c values",
       x = "Cross-subject variability (b)",
       y = "Within-subject variability (c)",
       fill = "SNR (dB)") +
  theme_minimal(base_size = 14)

result <- read.csv("signal_to_noise.csv")
mean(result$R2_marginal)
median(result$R2_marginal)
min(result$R2_marginal)
max(result$R2_marginal)
